Flowchart of Current Algorithm

### 2009-01-21 docs/flowchart/flowchart.tex
merged the CHANGES file into the flowchart "Algorithm Updates" section

\begin{itemize}
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form leads to better
numerical properties.  This is more like paper II.

@@

\item We discuss spherical geometries.

\item We add thermal diffusion.

\item We added the volume-discrepancy term to the velocity constraint
equation.
\end{itemize}

@@

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \er$) & cm~s$^{-1}$ \\
\hline
$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv (\overline{ \rho' \Ubt \cdot \er})$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}
%
\renewcommand{\arraystretch}{1.0}

We make use of the following shorthand notations in outlining the algorithm:
\begin{enumerate}

@@

In summary:
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
\end{array}\right.\nonumber
\end{equation}
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
\end{array}\right.
\end{equation}
Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.

>>>>

\begin{itemize}

\item We have added thermal diffusion with {\tt use\_thermal\_diffusion = T}.

\item We discuss spherical geometries, in paricular:
\begin{itemize}
\item We are working on {\bf Advect Base}, {\bf React Base}, and 
{\bf Correct Base} details.
\item We are working on $\etarho$ details.
\item We compute the full state T on cell-centers after the advection-diffusion step 
from $T=T(\rho,X_k,p_0)$ by setting {\tt use\_tfromp = T} rather than $T=T(\rho,h,X_k)$.
\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
update to instead be $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
\end{itemize}

\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form, $\rho h$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.

@@

\item We added the volume-discrepancy term to the velocity constraint equation.

\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff} = T.

\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.

\item We have added the flag, {\tt full\_rhoh0\_evolution}, when false, sets
$(\rho h)_0 = (\overline{\rho h})^n$ before {\bf Advect Base}.  {\bf React Base}
will not be called.  Note that we do not have a {\bf Correct Base} or 
{\bf Diffuse Base} for enthalpy.

\item We have given the advection scheme a face-lift, with some nifty options like
corner coupling and various PPM limiters.

\item We now update the density and enthalpy as full states with full-state fluxes, 
not as perturbational quantities.

\end{itemize}

@@

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \er$) & cm~s$^{-1}$ \\
\hline
$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv (\overline{ \rho' \Ubt \cdot \er})$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}
%
\renewcommand{\arraystretch}{1.0}
Here are some charts demonstrating what the superscript notation above scalars implies:
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
\end{array}\right.\nonumber
\end{equation}
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
\end{array}\right.\nonumber
\end{equation}
Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.\\ \\
We make use of the following shorthand notations in outlining the algorithm:
\begin{enumerate}

@@
